Our Title

### 2015-04-30 draft.tex
Merge remote-tracking branch 'origin/julie' into dev

\begin{itemize}
  \item Disrupt ISIL cash flow
  \item Rollback ISIL territorial gains
  \item Restrict ISIL access to nuclear weapons experts
\end{itemize}

@@

New report by Financial Action Task Force (an international body) finds that it is unclear whether ISIL’ revenue collection will be sustainable over time, and that ISIL will have to gain more territory to remain financially viable \cite{TheEditorialBoard2015,Report2015}.

ISIL employs an estimated 30,000 fighters, including 19,000 foreign fighters from 90 different countries, paying them roughly \$350-\$500 a month. This comes out to \$10 million per month. ISIL also tries to bolster its appeal by providing public services, including electricity, to people in its territory.

US strategy is to:

@@

In the current oil market, diminished profits from selling oil may make ISIL more dependent on external donations. As need for this arises, it is important to focus now on these external sources. Good work has been done in Saudi Arabia and UAE, but Kuwait and Qatar remain permissive jurisdictions. More needs to be done by these states to enforce their newly passed laws against financing terrorism, and US can help them do this through State Department, other training. Important to note that these Gulf states have an interest in stopping ISIL – ISIL has periodically targeted Qatar and its Ministry of Foreign Affairs.

Extortion and illicit taxation are large, internally sustainable methods of funding for ISIL. Before capturing Mosul, ISIL was already earning \$12 million a month in the city from these methods. A sophisticated ISIL taxation system has been created in on the main highway between Jordan and Baghdad that replaces the government’s import tax by charging a reduced rate for goods transported into the capital. ISIL does this throughout Iraq and Syria, and ensures its survivability by gaining tax income while charging low rates to Sunni truckers and other Sunnis to curry favor. Truly eroding this source of funding means breaking ISIL’ hold over the territory. This is highly unlikely to happen without sophisticated ground forces following up on the airstrikes.

@@

Disabling ISIL oil revenue is more complicated than it might seem. ISIL has been focused on establishing a durable internal market for oil, ensuring they have a reliable source of fuel for its fleets and also establishing dependence from the local population on its cheap oil. Recent airstrikes have targeted the oil at its source, but might be better served by targeting means of transportation and sale. Syrian rebels have pointed to this as a major error in targeting, as it disrupts oil supply to power generators and local facilities, allowing ISIL to pass the blame on to the US (who they call \enquote{the crusaders}). One difficulty in targeting the means of transportation and sale is that oil is transported in tanker trucks and oil drums, and most of the transactions are in cash, making them hard to track. One policy priority should be providing large quantities of diesel fuel and oil for generators in opposition areas of northern Syria [1].
A final recommendation is to put more intelligence resources toward detecting ISIL fundraising efforts through modern communication networks (social media) \cite{Report2015}.

@@

Share practical information and intelligence at an international level, both spontaneously and on request, to effectively disrupt international financial flows \cite{Report2015}.

@@

Request countries to proactively identify individuals and entities for inclusion in the UN Al Qaida Sanctions Committee list \cite{Report2015}.

@@

\subsection{Restrict ISIL access to nuclear weapons experts}

@@

@@

We should recommend that DoS is continuing with these initiatives, and that it possible increase funding for employment and research opportunities for nuclear scientists.

In terms of the second threat – that nuclear scientists aid ISIL for ideological reasons – is probably a lesser threat but more difficult to deal with. The steps that must be taken include a number of initiative to degrade ISIL's image as a leader in the Muslim world. Given the length of this report I'm thinking we may not want to go into this at length. Let me know if we need more research on this.

@@

ISIL could attempt to purchase or steal a nuclear weapon, and so the nations which maintain a stockpile of nuclear weapons must remain under scrutiny. Nations of the highest interest that have nuclear weapons are Pakistan, North Korea, India, Turkey, Israel, and Russia. Israel has not been confirmed to have nuclear weapons, but it is highly suspect of having them and Turkey has US tactical nuclear weapons stationed within them (NATO nuclear weapon sharing state). Both of these would unlikely have their weapons stolen as those in Turkey are under US protection and if Israel has them, they have been able to hide them from the world so it is unlikely a terrorist organization with much fewer intelligence capabilities would find them much less penetrate Israel’s security. Of the greatest concern is the possibility that any of these nuclear-capable states would willingly sell a physics package to ISIL.

The purchasing of nuclear materials, while could be done through a nation, is much more likely to be done through individuals. The IAEA keeps the only collection of nation confirmed information on incidents of illicit trafficking and other unauthorized activities involving nuclear and other radioactive material called the ITDB. As of 2006 over 1000 incidents of illicit trafficking and other unauthorized activities involving nuclear and other radioactive material has been reported to the IAEA with about 25\% specifically involving nuclear materials \cite{Iaea2007}. A majority of these cases involve highly enriched uranium (no plutonium). One such incident is in 1994 when an individual stole 2.972 kg of highly enriched uranium from a nuclear factory \cite{Iaea2007}. See \autoref{fig:seizures} for more information on the location and size of notable seizures.

@@

Most of these instances are very small amounts (grams) but there have been several involving one to three kilograms. No individual case has been found with enough material to make a nuclear bomb, but with multiple cases it would have been possible to accumulate enough special nuclear material to create a nuclear weapon.

The non-proliferation treaty (NPT) has strengths that assist with prevention of these nations selling or having their weapons stolen. NPT has a multitude of safeguards in place, including accounting for the flow of materials in and out of nuclear facilities, physical security, and containment and surveillance.However, while the NPT’s strength has been in preventing current nations from using existing energy facilities to making off the books weapon grade materials (these would be easier to sell), Iran, Iraq, and North Korea were able to set up separate facilities for creating weapons grade material. This could result in counties with the capabilities being able to create weapons grade material outside the safeguards of the NPT. However, it should be noted, that while these facilities were set up without knowledge, they were discovered in Iraq and North Korea rather quickly.

\subsubsection{RDD}

A portion of these materials which can be integrated into a RDD are naturally occurring in nature, and are intrinsically radioactive, while others are solely limited in availability to radiological work sites such as reactor facilities, isotope production, research and universities, weapons production facilities, and other commercial and industrial facilities.

@@

Identify possible and likely sources of the top 10 RDD materials identified in section 2.2 that ISIL might purchase or steal. Also identify possible sources of these materials within territory that ISIL controls (such as hospitals and universities). Discuss any current export controls on these materials, and how they might inhibit the acquisition of significant quantities of materials. Discuss the orphan source program, or any other international institutions which regulate the creation and distribution of radioactive sources. (To be added by 4/19; possible add map of sources?)

@@

Up to this point, the main focus of this policy has been the prevention of ISIL from developing, stealing, or buying a nuclear weapon or RDD.  However, it is necessary to develop a contingency in the unlikely event of failure in which ISIL obtains a weapon.  Unfortunately, the US cannot use the same nuclear strategies it has in the past to deal with a nuclear ISIL.  The main focus of US nuclear policy since the collapse of the Soviet Union has been non-proliferation.  There has not been as strong of a focus on deterring attacks from an adversary that already has nuclear capabilities \cite{Bracken2013}. However, even before the fall of the Soviet Union, deterrent strategies, such as counterforce, counter value, and mutually assured destruction, were tailored to deter an enemy state, not an extremist group.  Deterrence can only happen if the adversarial group/state thinks that the punishment received from the US will outweigh the benefits of detonating a nuclear weapon \cite{kaplan1991wizards}.  It is unlikely that the US will ever be able to make a threat that can deter ISIL.  Unlike the Soviet Union and other nuclear states, ISIL is made up entirely of militants who will die in defense of their religion and their caliph.  Death is not something that members of ISIL, or any other terrorist group, fear; it is something they celebrate.  They believe they will be rewarded in the next life for what their actions on earth.  Even the threat of bombing by the US and death will not deter them.  If they have reached the point of weapon acquisition, they have already ignored the United States’ threats of massive retaliation and will likely continue to ignore any future threats.  For this reason, deterrence will be ineffective if ISIL acquires a bomb, so the US must focus on other methods of stopping ISIL from detonating the weapon \cite{Sanderson2015}.

@@

The first step the US can take after bomb acquisition by ISIL is to cut them off at their delivery method.  Generally, for nuclear powers, delivery methods are described by the nuclear triad: ground based missiles, such as ICBMs, strategic air bombers, and submarine-launched missiles.  However, as North Korea has shown, ground based missiles are not easily implemented.  It is highly unlikely that ISIL will be able to construct a long-range missile that can carry a nuclear weapon or a submarine capable of launching it.  North Korea has spent years trying to build ballistic missiles capable of delivering nuclear weapons over long ranges, and it is still questionable if they are capable of medium range distances \cite{Gladstone2015}.  It is also highly unlikely that any country will sell long-range missiles to ISIL.  As mentioned in \autoref{sec:global_rels} (\enquote{Global Relationships}), even though North Korea, Iran, and Pakistan are the most likely to sell weapons to ISIL, it is still a low probability.  Also, none of these countries have ICBMs. North Korea and Pakistan do have short-range and some medium-range missiles \cite{Gladstone2015,NuclearThreatInitiative2014}.  Though, it is unlikely that these would be sold to ISIL, these countries should be heavily monitored in the event of bomb acquisition.  Submarine capabilities seem highly unlikely as well since they are more difficult to develop than land based missiles and the land that ISIL currently occupies is land locked.  The third part of the nuclear triad is long-range bombers.  However, few countries have long-range bombers and, of those countries, it is highly improbable that any of them would sell or give ISIL a bomber.  Therefore, the most likely method of delivery would be to ship or carry the bomb to the target depending on the location.

The next step the US can take is to predict the targets and intercept the bomb in transit to those targets.  It is unlikely that ISIL would use a nuclear weapon or RDD on cities in Syria or Iraq that they do not occupy since that would make the city and their resources unusable to them.  The fallout afterwards would also be close enough to hurt them as well.  It is more likely they will attack one of the other 61 countries in the Anti-ISIL Coalition \cite{Wordsworth2015}.  As mentioned earlier in this policy, it can be assumed that ISIL will follow a counter value strategy versus counterforce.  They are more likely to attack one of the population centers in one of these countries.  A major US city would initially seem a primary target, however, due to the geographical distance from ISIL, it seems unlikely.  Since missile delivery is highly improbable, it would have to be shipped to the US.  However, there is no direct shipment method from ISIL to the US.  So, the bomb would be shipped through multiple ports and pass through multiple nuclear detector systems before reaching a US port.  However, this possibility should not be ignored.  For this reason, if it is even suspected that ISIL has acquired a nuclear weapon, security at all US ports should be increased.  Portal detectors should be used at all major ports and border crossings and at as many smaller ports as possible.  There is also the possibility of the attack of European countries.   However, as with the US, it is unlikely that ISIL will be able to ship it directly to any of the countries in Europe without it being detected.  However, another reason that European countries are not likely targets is due to the ideology behind ISIL.  ISIL’s main targets are currently other Muslims who do not hold their extremist beliefs.  They are slightly more tolerant of Christians and Jews than \enquote{apostate} Muslims.  They even let Christians remain in Baghdad \cite{Wood2015}.  So, ISIL is more likely to attack another Muslim nation, especially one that is geographically close.  Due to their close proximity and involvement in the Anti-ISIL coalition, Saudi Arabia, Jordan, Turkey and Lebanon are likely targets.  Iran is also another likely target due to its large Shia population.  Due to US-Iranian relations and US sanctions on Iran, it is unlikely that the US and Iran will be able to work together on stopping a bomb from crossing over the Iranian border.  However, for the other countries, the US can assist in monitoring their borders and possibly supply these countries with more advanced nuclear detection equipment.  With the US and other countries in the Anti-ISIL Coalition focusing on these countries and their borders, it may be possible to stop the bomb from being delivered to its intended target.

>>>>

\begin{itemize}
  \item Disrupt ISIL cash flow
  \item Rollback ISIL territorial gains
  \item Counteract ISIL's draw to foreign fighters
\end{itemize}

@@

A new report by Financial Action Task Force (an international body) finds that it is unclear whether ISIL’ revenue collection will be sustainable over time, and that ISIL will have to gain more territory to remain financially viable \cite{TheEditorialBoard2015,Report2015}.

ISIL employs an estimated 30,000 fighters, including 19,000 foreign fighters from 90 different countries, paying them roughly \$350-\$500 a month. This leads to a total payroll cost of \$10 million per month. ISIL also tries to bolster its appeal by providing public services, including electricity, to people in its territory.

Current US strategy is to:

@@

In the current oil market, diminished profits from selling oil may make ISIL more dependent on external donations. As need for this arises, it is important to focus on these external sources. Saudi Arabia and UAE are becoming more effective at preventing funding of terrorist orgnaizations by their citizens, but Kuwait and Qatar remain permissive jurisdictions. More needs to be done by these states to enforce their newly passed laws against financing terrorism, and the US can help them do this through State Department and other training. It is important to note that these Gulf states have an interest in stopping ISIL – ISIL has periodically targeted Qatar and its Ministry of Foreign Affairs. \subsubsection{???????? THIS PARAGRAPH NEEDS SOURCES   ?????????}

Extortion and illicit taxation are large, internally sustainable methods of funding for ISIL. Before capturing Mosul, ISIL was already earning \$12 million a month in the city from these methods. A sophisticated ISIL taxation system has been created on the main highway between Jordan and Baghdad that replaces the government’s import tax by charging a reduced rate for goods transported into the capital. ISIL does this throughout Iraq and Syria, and ensures its survivability by gaining tax income while charging low rates to Sunni truckers and other Sunnis to curry favor. Truly eroding this source of funding means breaking ISIL’ hold over the territory. This is highly unlikely to happen without sophisticated ground forces following up on the airstrikes.

@@

Disabling ISIL oil revenue is more complicated than it might seem. ISIL has been focused on establishing a durable internal market for oil, ensuring they have a reliable source of fuel for its fleets and also establishing dependence from the local population on its cheap oil. Recent airstrikes have targeted the oil at its source, but might be better served by targeting means of transportation and sale. Syrian rebels have pointed to this as a major error in targeting, as it disrupts oil supply to power generators and local facilities, allowing ISIL to pass the blame on to the US (who they call \enquote{the crusaders}) when the electricity system failes. One difficulty in targeting the means of transportation and sale is that oil is transported in tanker trucks and oil drums, and most of the transactions are in cash, making them hard to track. One policy priority should be providing large quantities of diesel fuel and oil for generators in opposition areas of northern Syria [1].
A final recommendation is to put more intelligence resources toward detecting ISIL fundraising efforts through modern communication networks (social media) \cite{Report2015}.

@@

The US should share practical information and intelligence at an international level, both spontaneously and on request, to effectively disrupt international financial flows \cite{Report2015}.

@@

The US should request countries to proactively identify individuals and entities for inclusion in the UN Al Qaida Sanctions Committee list \cite{Report2015}.

 \subsubsection{???????? Say more about this, what is this list??   ?????????}

@@

\subsection{Counteract ISIL's draw to foreign fighters.}

ISIL has shown itself prolific at recruiting foreign fighters, and the US must try to stem this flow in its entireity, but alos to prevent nuclear scientists from joining ISIL's causes.

@@

\subsubsection{???????? not sure the first paragraph of this quote is necessary, unless your saying that we should spend comprable amounts of money now. also, you can probably just intergrate the points of the second paragraph into your own paragraph, no real need to pull it out as a block quote  ?????????}

@@

We should recommend that DoS is continuing with these initiatives, and if possible increase funding for employment and research opportunities for nuclear scientists. The second threat – that nuclear scientists will aid ISIL for ideological reasons – is of lesser probability, \subsubsection{???????? source?? ?????????} but more difficult to deal with. The steps that must be taken include a number of initiative to degrade ISIL's image as a leader in the Muslim world, which is part of the larger fight against ISIL as a whole, and is not included in depth in this report.

@@

ISIL could attempt to purchase or steal a nuclear weapon, and so the nations which maintain a stockpile of nuclear weapons must remain under scrutiny. Nations of the highest interest that have nuclear weapons are Pakistan, North Korea, India, Turkey, Israel, and Russia. Israel has not been confirmed to have nuclear weapons, but it is widely believed that it has nuclear capabilites. Turkey is a NATO nuclear weapon sharing state and has US tactical nuclear weapons stationed within its borders. However, it is unlikely that ISIL would be able to steal a nuclear weapon from either of these states, as those in Turkey are under US protection and Israel has proven adept at hiding their arsenal. Of greatest concern is the possibility that any of these nuclear-capable states, North Korea in particular, would willingly sell a physics package to ISIL.

The purchasing of nuclear materials, while it could be done on a national level, is much more likely to occur through individuals. The only collection of nation confirmed information on incidents of illicit trafficking and other unauthorized activities involving nuclear and other radioactive material is called the ITDB \subsubsection{???????? what does this acronym stand for?  ?????????} and is maintained by the IAEA. As of 2006, over 1000 incidents of illicit trafficking and other unauthorized activities involving nuclear and other radioactive material has been reported to the IAEA with about 25\% specifically involving nuclear materials \cite{Iaea2007}. A majority of these cases involve highly enriched uranium (no plutonium). One such incident occurred in 1994 when an individual stole 2.972 kg of highly enriched uranium from a nuclear factory \cite{Iaea2007}. See \autoref{fig:seizures} for more information on the location and size of notable seizures.

@@

Most of these instances are very small amounts (grams) but there have been several involving one to three kilograms. No individual case has been found with enough material to make a nuclear bomb, but muiltiple thefts could allow ISIL to accumulate enough special nuclear material to create a nuclear weapon.

The non-proliferation treaty (NPT) assists with prevention of these nations selling or having their weapons stolen. The NPT has a multitude of safeguards in place, including accounting for the flow of materials in and out of nuclear facilities, physical security, and containment and surveillance.However, while the NPT’s has been effective in preventing nations from using existing energy facilities to make off the books weapon grade materials (these would be easier to sell), Iran, Iraq, and North Korea were able to set up separate facilities for creating weapons grade material. While these facilities were set up without international knowledge, they were discovered in Iraq and North Korea rather quickly.

\subsubsection{RDD}
 Some of the materials required to build an RDD are naturally occurring elements that are naturally radioactive. Other components are solely availabile to radiological work sites such as reactor facilities, isotope production, research and universities, weapons production facilities, and other commercial and industrial facilities.

@@

\subsubsection{???????? Identify possible and likely sources of the top 10 RDD materials identified in section 2.2 that ISIL might purchase or steal. Also identify possible sources of these materials within territory that ISIL controls (such as hospitals and universities). Discuss any current export controls on these materials, and how they might inhibit the acquisition of significant quantities of materials. Discuss the orphan source program, or any other international institutions which regulate the creation and distribution of radioactive sources. (To be added by 4/19; possible add map of sources?) ?????????????}

@@

Up to this point, the main focus of this policy has been the prevention of ISIL from developing, stealing, or buying a nuclear weapon or RDD.  However, it is necessary to develop a contingency in the unlikely event of failure in which ISIL obtains a weapon.  Unfortunately, the US cannot use the same nuclear strategies it has in the past to deal with a nuclear ISIL.  The main focus of US nuclear policy since the collapse of the Soviet Union has been non-proliferation.  There has not been as strong of a focus on deterring attacks from an adversary that already has nuclear capabilities \cite{Bracken2013}. However, even before the fall of the Soviet Union, deterrent strategies, such as counterforce, counter value, and mutually assured destruction, were tailored to deter an enemy state, not an extremist group.  Deterrence can only happen if the adversarial group/state thinks that the punishment received from the US will outweigh the benefits of detonating a nuclear weapon \cite{kaplan1991wizards}.  It is unlikely that the US will ever be able to make a threat that can deter ISIL.  Unlike the Soviet Union and other nuclear states, ISIL is made up entirely of militants who will die in defense of their religion and their caliph.  Death is not something that members of ISIL, or any other terrorist group, fear; it is something they celebrate.  They believe they will be rewarded in the next life for what actions on earth.  Even the threat of bombing by the US and death will not deter them.  If they have reached the point of weapon acquisition, they have already ignored the United States’ threats of massive retaliation and will likely continue to ignore any future threats.  For this reason, deterrence will be ineffective if ISIL acquires a bomb, so the US must focus on other methods of stopping ISIL from detonating the weapon \cite{Sanderson2015}.

@@

The first step the US can take after bomb acquisition by ISIL is to cut them off at their delivery method.  Generally, for nuclear powers, delivery methods are described by the nuclear triad: ground based missiles, such as ICBMs, strategic air bombers, and submarine-launched missiles.  However, as North Korea has shown, ground based missiles are not easily implemented.  It is highly unlikely that ISIL will be able to construct a long-range missile that can carry a nuclear weapon or a submarine capable of launching it.  North Korea has spent years trying to build ballistic missiles capable of delivering nuclear weapons over long ranges, and it is still questionable if they are capable of medium range distances \cite{Gladstone2015}.  It is also highly unlikely that any country will sell long-range missiles to ISIL.  As mentioned in \autoref{sec:global_rels} (\enquote{Global Relationships}), even though some nations may sell weapons to ISIL, it is still a low probability.  Also, none of these countries have ICBMs. North Korea and Pakistan do have short-range and some medium-range missiles \cite{Gladstone2015,NuclearThreatInitiative2014}.  Though, it is unlikely that these would be sold to ISIL, these countries should be heavily monitored in the event of bomb acquisition.  Submarine capabilities seem highly unlikely as well since they are more difficult to develop than land based missiles and the land that ISIL currently occupies is land locked.  The third part of the nuclear triad is long-range bombers.  However, few countries have long-range bombers and, of those countries, it is highly improbable that any of them would sell or give ISIL a bomber.  Therefore, the most likely method of delivery would be to ship or carry the bomb to the target depending on the location.

The next step the US can take is to predict the targets and intercept the bomb in transit to those targets.  It is unlikely that ISIL would use a nuclear weapon or RDD on cities in Syria or Iraq that they do not occupy since that would make the city and their resources unusable to them.  The fallout afterwards would also be close enough to hurt them as well.  It is more likely they will attack one of the other 61 countries in the Anti-ISIL Coalition \cite{Wordsworth2015}.  As mentioned earlier in this policy, it can be assumed that ISIL will follow a counter value strategy versus counterforce.  They are more likely to attack one of the population centers in one of these countries.  A major US city would initially seem a primary target, however, due to the geographical distance from ISIL, it seems unlikely.  Since missile delivery is highly improbable, it would have to be shipped to the US.  However, there is no direct shipment method from ISIL to the US.  So, the bomb would be shipped through multiple ports and pass through multiple nuclear detector systems before reaching a US port.  However, this possibility should not be ignored.  For this reason, if it is even suspected that ISIL has acquired a nuclear weapon, security at all US ports should be increased.  Portal detectors should be used at all major ports and border crossings and at as many smaller ports as possible.  There is also the possibility of the attack of European countries.   However, as with the US, it is unlikely that ISIL will be able to ship it directly to any of the countries in Europe without it being detected.  However, another reason that European countries are not likely targets is due to the ideology behind ISIL.  ISIL’s main targets are currently other Muslims who do not hold their extremist beliefs.  They are slightly more tolerant of Christians and Jews than \enquote{apostate} Muslims. \cite{Wood2015}  ISIL is therefore more likely to attack another Muslim nation, especially one that is geographically close.  Due to their close proximity and involvement in the Anti-ISIL coalition, Saudi Arabia, Jordan, Turkey and Lebanon are likely targets.  Iran is also another likely target due to its large Shia population.  Due to US-Iranian relations and US sanctions on Iran, it is unlikely that the US and Iran will be able to work together on stopping a bomb from crossing over the Iranian border.  However, for the other countries, the US can assist in monitoring their borders and possibly supply these countries with more advanced nuclear detection equipment.  With the US and other countries in the Anti-ISIL Coalition focusing on these countries and their borders, it may be possible to stop the bomb from being delivered to its intended target.
